Towards an ecology of vagueness

### 2017-11-11 book_chapter/paper.tex
Re-wrote conclusions.

\section{Conclusions}
\label{sec:conclusions}
\todo[inline]{Extend/Rewrite}
We tried to show that, if we pragmatically acknowledge the complex and dynamic environment that surrounds language learning and use, we find many plausible sources of uncertainty that can lead to vague signal use.
We supported this claim by giving an overview of recent research that tries to accommodate for vagueness in the context of sim-max games, and by discussing, in more general terms, how these considerations are relevant for reference games, and in the broader context of language as a multiplicity of language-games.
Furthermore, we argued that if we broaden our notion of rationality beyond mere maximization of expected utility, we also find reasons to welcome the positive aspects of these processes and to see their acceptance as a perfectly rational choice.
Irrational would be to ignore the finite nature of our existence and the constraints that condition language learning and use.
Our challenge as researchers is of course not only to draw attention to the possibilities, but also to characterize which particular constraints give rise to vague signal use and how.

The investigation doesn't end here, then.
The research discussed here suggests that although vagueness can be beneficial, that is the case only in moderation: too much can lead to babbling equilibria where no information is transmitted.
Thus, for agents capable of adjusting the level of vagueness in their strategies, questions remain regarding how this can be modeled and what factors influence the amount of vagueness that is optimal for a given situation.
Other challenges include how to incorporate into our models the intuitions about the importance of speed of convergence, adaptability, and the promotion of group homogeneity.
Finally, perhaps the biggest of them all is how to be able to focus on one language-game without losing sight of the ecosystem of games that need to be taken into account as well.
Whatever solutions we find to these challenges, we believe that a notion of rationality cannot be strongly tied to mere utility maximization in a particular game, but has to consider an agents' need to optimize his behavior to navigate the plurality of uncertainties that besiege him.
We need a clear notion of ecological rationality in order to see vagueness in the right light.

>>>>

\section{Conclusions}
\label{sec:conclusions}
Vagueness presents a challenge to both procedural and instrumental pictures of language and rationality alike.
In the context of game-theoretical models of language, this takes the form of a question about evolution: how can vagueness persist if leads to demonstrably less efficient communication?
Most existing proposals in the literature attempt to explain this by considering agents with some degree of bounded rationality.
Despite the many ways in which our rationality is inevitably limited, the argument for the pervasiveness of vagueness in natural language would be much strengthened if there were also associated advantages.
In this paper, we argued that finding those might require us to go beyond a local notion of rationality, as two of the proposals reviewed here~\parencite{oconnor_evolution_2014, franke_vagueness_2017} suggest.
We advocate moving towards an ecological approach, studying vagueness in more heterogenenous ecosystems.
Language is part of a very complex system~\parencite{beckner_language_2009} involving many components that interact in often unpredictable ways.
An ecology of vagueness would involve considering models where different populations can evolve and interact with each other, or where different language games can be played between the individuals, or where the environment is uncertain and changing, all of the above, or anything else that more closely approximates the real context of language evolution.

In light of this picture, we proposed two variants of a concrete multi-population signaling model to test the hypothesis~\parencite{franke_vagueness_2017} that certain features of imprecise imitation, like promoting faster convergence and regularity, could prove beneficial against full precision.
Analysing these models did not provide us with a clear-cut answer, and revealed that the story is much more nuanced than initially expected.
In a variant where populations with and without imprecision interact tightly, although precision always has the upper hand, vagueness seems to take a long time to be weeded out.
When we let populations interact more loosely, we see a more complex pattern of outcomes.
These include scenarios where imprecise imitation dominates over full precision, showing that strategies with vagueness can actually, under certain circumstances, be more successful.
Bringing several populations together in a more complex ecosystem thus allowed us to not only spell out and test the original intuition, but also learn about new unforeseen effects.
These models thus serve as an example of how moving to a more global perspective on rationality can allow us to achieve a more detailed awareness of the complex interactions that might be involved in sustaining vagueness in our language.
They can be seen, we believe, as a first step towards an ecology of vagueness.
